---
title: "P8105 HW3"
author: "Hansheng Zhang"
date: 2021-10-20
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(p8105.datasets)

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

Exploration of instacart dataset

``` {r load_data}
library(p8105.datasets)
data("instacart")
data("brfss_smart2010")
```

The `instacart` dataset has `r nrow(instacart)` observations of `r ncol(instacart)` variables. The variable names are `r names(instacart)`. Variables that are specific to the items include `product_id`, `department`, `product_name`, and `aisle`. Vvariables that are specific to the consumer include `reordered`, `user_id`, `days_since_prior_order`, and `order_dow`. This dataset has `r instacart %>% distinct(user_id) %>% count()` unique customers.

```{r aisle_count}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are `r instacart %>% distinct(aisle) %>% count()` aisles.
The aisles items are ordered from the most are:

1. fresh vegetables  
1. fresh fruits  
1. packaged vegetable fruits  
1. yogurt  
1. packaged cheese  

Plot that shows number of items ordered in each aisle

```{r aisles_plot_instacart}
instacart %>% 
  group_by(aisle) %>% 
  summarize(item_count = n()) %>% 
  filter(item_count > 10000) %>% 
  ggplot(aes(x = reorder(aisle, item_count), y = item_count)) +
  geom_bar(stat = "identity") +
  coord_flip() + 
  labs(
    title = "Number of items ordered in each aisle (>1000)",
    x = "Aisle name",
    y = "Items ordered",
    caption = "Data from the instacart dataset"
  )
```

This table shows the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” and the number of times each item is ordered.

```{r instacart_table1}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r instacart_table2}
instacart %>% 
  mutate(
  order_dow = recode(order_dow,
                       "0" = "Sunday",
                       "1" = "Monday",
                       "2" = "Tuesday",
                       "3" = "Wednesday",
                       "4" = "Thursday",
                       "5" = "Friday",
                       "6" = "Saturday")
  ) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  select(product_name, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday,
         everything())
```


# Problem 2

Data cleaning and exploration of BRFSS data set

```{r}
brfss_df = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
                      ordered = TRUE
                      )
    ) %>% 
    rename(
    state = locationabbr,
    location = locationdesc
    )
```


```{r}
brfss_2002 =
  brfss_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  summarize(num_location = n()) %>% 
  filter(num_location >= 7) %>% 
  pull(state)
brfss_2010 =
  brfss_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  summarize(num_location = n()) %>% 
  filter(num_location >= 7) %>% 
  pull(state)
```

In 2002, states that were observed at 7 or more locations are `r brfss_2002`. In 2010, states that were observed at 7 or more locations are `r brfss_2010`.

Plot of the average data value over time within a state

```{r}
excellent_df <- 
  brfss_df %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  summarize(
    mean_data_value = mean(data_value, na.rm = TRUE)
  )
excellent_df %>% 
  ggplot(aes(x = year, y = mean_data_value, color = state)) +
  geom_line(alpha = 0.5) +
  labs(
    y = "average data value",
    title = "Average Value over Time within Each States",
    caption = "Data from BRFSS data set"
  ) +
  theme(
    plot.caption = element_text(size = 5),
    legend.key.size = unit(0.4, "cm"),
    legend.text = element_text(size = 5)
  ) +
  guides(color = guide_legend(nrow = 4))
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State

```{r}
brfss_df %>% 
  filter(year %in% c(2006, 2010), state == "NY") %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .3) +
  facet_grid(. ~ year) +
  labs(y = "density of data value",
       x = "data value",
    title = "Distribution of Data Value for Reponse among New York State in 2006 vs. 2010",
    caption = "Data from BRFSS data set")
```


# Problem 3

Tidy and clean accelerometer data and then create a table and a plot. 

```{r,warning=FALSE}
accel_data = read_csv("./Data/accel_data.csv") %>% 
    pivot_longer(
        activity.1:activity.1440,
        names_to = "activity",
        values_to = "activity_count") %>% 
    mutate(
        day_type = case_when(
            day %in% c("Saturday", "Sunday") ~ "weekend",
            TRUE ~ "weekday"),
        day = fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))
```

There are total `r nrow(accel_data)` observations and `r ncol(accel_data)` variables. The variables in this data set are `r names(accel_data)`.

```{r activity_count_table}
accel_data %>% 
    group_by(week, day) %>% 
    summarize(day_total = sum(activity_count)) %>% 
    mutate(day_total = round(day_total, digits = 0)) %>% 
    pivot_wider(
        names_from = day,
        values_from = day_total) %>% 
    rename("Week" = week) %>% 
    knitr::kable()
```

No significant activity trends appear. However, the last two weekends showed much lower activity counts than in previous weeks and weekends. Furthermore, the middle of the week generally showed less activity counts than other days of the week.  


Time plot of all the activity metrics across each day of the week by the minute.  

```{r activity_count_plot}
accel_data %>% 
    mutate(min_of_day = as.integer(str_sub(activity, 10, -1))) %>% 
    ggplot(aes(x = min_of_day, y = activity_count, color = day)) +
    geom_line() +
    scale_x_continuous(
        breaks = c(0, 360, 720, 1080, 1440),
        labels = c("00:00", "06:00", "12:00", "18:00", "23:59")) +
    labs(
        title = "Daily activity",
        x = "Time of Day",
        y = "Level of Activity") +
    viridis::scale_color_viridis(
        name = "Day of the Week", 
        discrete = TRUE)
```

The patient had increased activity on Sunday around 11:00, and on Friday and Saturday night around 20:00 to 22:00. 

